Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-01-31 main.tex
recursive estimate

\begin{theorem}\label{theorem:globalestimate:gradient}
    Let $\calT$ be a face-connected $n$-dimensional finite triangulation
    and that the domain $\Omega$ is the interior of the underlying set of $\calT$. 
    Let $1 \leq p,q \leq \infty$ with $1 = 1/p + 1/q$,
    and suppose that $u, w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$.

    Suppose that $n$-simplices are enumerated as $T_0, T_1, \dots, T_M$,
    and that with each $n$-simplex $T_{m} \in \calT$ 
    we have a sequence of indices 
    \begin{gather*}
        0 = \imath(m,0), \; \imath(m,1), \; \dots, \; \imath(m,L_m) = m
    \end{gather*}
    such that for each $0 \leq m \leq M$ we have estimates of the following form:
    \begin{align*}
        \| w \|_{L^{p}(T_{0})} &\leq A_{m,0} \| \nabla u \|_{L^{p}(T_{0})}
        ,
    \end{align*}
    \begin{align*}
        \| w \|_{ L^{p}(T_{\imath(m,\ell)}) }
        &\leq  
        A_{m,\imath(m,\ell)}
        \| w \|_{ L^{p}(T_{\imath(m,\ell-1)}) }
        \\&\qquad 
        +
        B'_{m,\imath(m,\ell)}
        \| \nabla u \|_{ L^{p}(T_{\imath(m,\ell)}) }
        + 
        B''_{m,\imath(m,\ell)}
        \| \nabla u \|_{ L^{p}(T_{\imath(m,\ell-1)}) }
        ,
        \quad 1 \leq \ell \leq L_m
        .
    \end{align*}
    Then 
    \begin{gather*}
        \| w \|_{ L^{p}(\Omega) }
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{L_m} D_{m,\ell}^{q} \right)^{\frac p q}
        \right)^{\frac 1 p}
        \| \nabla u \|_{ L^{p}(\Omega) }
        ,
    \end{gather*}
    where 
    \begin{gather*}
        D_{m,L_m} = A_{m,m},
        \\
        D_{m,\ell} 
        = 
        \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) 
        \left( B_{m,\imath(m,\ell+1)} A_{m,\imath(m,\ell)} + A'_{m,\imath(m,\ell+1)} \right) 
        ,
        \qquad 
        1 \leq \ell \leq L_{m}-1
        .
    \end{align*}
    % TODO: complete here. We need a spanning tree. 
\end{theorem}
\begin{proof}
    Unwrapping the recursion for the norm over the $m$-th simplex, we get a total expression of the form 
    \begin{align*}%\label{math:fullrecursivesum}
        \begin{split}
            \| w \|_{L^{p}(T_{m})} 
            &
            \leq 
            \sum_{\ell=0}^{L_m}   \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+1)} \right) A_ {m,\imath(m,\ell  )} \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&\qquad 
            +
            \sum_{\ell=0}^{L_m-1} \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) A'_{m,\imath(m,\ell+1)} \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&
            = 
            A_{m,m} \| \nabla u \|_{L^{p}(T_{\imath(m,L_m)})}
            \\&\qquad 
            +
            \sum_{\ell=0}^{L_m-1} 
            \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) 
            \left( B_{m,\imath(m,\ell+1)} A_{m,\imath(m,\ell)} + A'_{m,\imath(m,\ell+1)} \right) 
            \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&
            =:
            \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})}
            .
        \end{split}
    \end{align*}
    Here, $C_{m,\ell}$ is the coefficient of $\| \nabla u \|_{L^{p}(T_{\ell})}$, possibly zero, as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \begin{split}
            \| w \|_{L^{p}(\Omega)}^{p}
            &
            \leq 
            \sum_{m=0}^{M}
            \| w \|_{L^{p}(T_{m})}^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})} \right)^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \sum_{\ell'=0}^{M} \| \nabla u \|_{L^{p}(T_{\ell'})}^{p} 
            \leq 
            \left(
                \sum_{m=0}^{M}
                \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \right)
            \| \nabla u \|_{L^{p}(\Omega)}^{p} 
            ,
        \end{split}
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
\end{proof}

\begin{remark}\label{remark:fullrecursivesum}
    \color{red}
    We briefly outline how  to obtain an estimate for the Poincar\'e--Friedrichs constant from the recursive estimate in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
    The potential $w$ for the gradient $\nabla u$ satisfies 
    \begin{align*}
        \| w \|_{L^{p}(T_{0})} \leq A_{0} \| \nabla u \|_{L^{p}(T_{0})},
        \qquad 
        % \| w \|_{L^{p}(T_{m})} \leq A_{m} \| \nabla u \|_{L^{p}(T_{m} \cup T_{m-1})}                             + B_{m} \| w \|_{L^{p}(T_{m-1})},
        \| w \|_{L^{p}(T_{m})} \leq A_{m} \| \nabla u \|_{L^{p}(T_{m})} + A'_{m} \| \nabla u \|_{L^{p}(T_{m-1})} + B_{m} \| w \|_{L^{p}(T_{m-1})},
    \end{align*}
    where $A_{m}, A'_{m}, B_{m} \geq 0$ are numbers and $T_{0}, T_{1}, \dots, T_{m-1}, T_{m}$ is a face path.

    Unwrapping the recursion for the norm over the $m$-th simplex, we get a total expression of the form 
    \begin{align}\label{math:fullrecursivesum}
        \begin{split}
            \| w \|_{L^{p}(T_{m})} 
            &
            \leq 
            \sum_{\ell=0}^{m}   \left( B_{m,m} \cdots B_{m,\ell+1} \right) A_{m,\ell} \| \nabla u \|_{L^{p}(T_{m,\ell})}
            +
            \sum_{\ell=0}^{m-1} \left( B_{m,m} \cdots B_{m,\ell+2} \right) A'_{m,\ell+1} \| \nabla u \|_{L^{p}(T_{m,\ell})}
            \\&
            = 
            A_{m,m} \| \nabla u \|_{L^{p}(T_{m,m})}
            +
            \sum_{\ell=0}^{m-1} \left( B_{m,m} \cdots B_{m,\ell+2} \right) \left( B_{m,\ell+1} A_{m,\ell} + A'_{m,\ell+1} \right) \| \nabla u \|_{L^{p}(T_{m,\ell})}
            \\&
            =:
            \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})}
            .
        \end{split}
    \end{align}
    Here, $C_{m,\ell}$ is the coefficient of $\| \nabla u \|_{L^{p}(T_{\ell})}$, possibly zero, as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    By definition, $C_{m,0} = A_0$, and 
    \begin{align}
        C_{m,}
    \end{align}

    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \begin{split}
            \| w \|_{L^{p}(\Omega)}^{p}
            &
            \leq 
            \sum_{m=0}^{M}
            \| w \|_{L^{p}(T_{m})}^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})} \right)^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
            \sum_{\ell=0}^{M} 
            \| \nabla u \|_{L^{p}(T_{\ell})}^{p} 
            \leq 
            \left(
                \sum_{m=0}^{M}
                \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \right)
            \| \nabla u \|_{L^{p}(\Omega)}^{p} 
            ,
        \end{split}
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$.

    In other words, the Poincar\'e--Friedrichs inequality is estimated by first taking the $q$-norm of the coefficients in the full sum~\eqref{math:fullrecursivesum} and then taking $p$-norm of these intermediate sums.
\end{remark}

@@

    We now define, as in~\cite[Equations~{\char`\(}5.12{\char`\)} and~{\char`\(}5.14{\char`\)}]{ern2020stable} or~\cite[equations~{\char`\(}6.7{\char`\)} and~{\char`\(}6.9{\char`\)}]{Chaum_Voh_p_rob_3D_H_curl_24}, the field $\widetilde w_{m} \in W^{p}\Alt^{k}(T_{m})$ via 
    \begin{align} \label{equation:w''_ext_der}
        {\widetilde w}_{m} := u_{|T_m} + \Xi_{1}^{\ast}\left( (w_{m-1} - u)_{|\Makoto_{m-1} } \right).
    \end{align}
    One verifies that 
    \begin{align*}
        \trace_F {\widetilde w}_{m} 
        &
        = 
        \trace_F u_{|T_m} + \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \trace_F \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \trace_F u_{|T_m} + \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \trace_F \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} }
        \\&
        = 
        \trace_F w_{m-1|\Makoto_{m-1}}
        .
    \end{align*}
    Moreover, 
    \begin{align*}
        \cartan {\widetilde w}_{m} 
        &
        = 
        \cartan u_{|T_m} + \cartan \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \cartan \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \cartan u_{|T_m} + \Xi_{1}^{\ast} \cartan w_{m-1|\Makoto_{m-1} } - \Xi_{1}^{\ast} \cartan u_{|\Makoto_{m-1} } 
        = 
        \cartan u_{|T_m}
        .
    \end{align*}
    Setting $w_{m} := w_{m-1}$ over $X_{m-1}$ and $w_{m} := {\widetilde w}_{m} + w''_{m}$ over $T_{m}$, 
    we thus verify $w_{m} \in W^{p}\Alt^{k}(X_{m})$ with $\cartan w_{m} = \cartan u$ over $X_{m}$.

@@

\begin{remark}\label{remark:poincarefriedrichsestimate:exterior:contraction}
    We would like to point out an alternative application of Proposition~\ref{proposition:starreflection},
    using the second class of transformations. 
    Those transformations are a sequence of contractions.
    Here, we take inspiration from the topological observation that domains with shellable triangulations are contractible. 
    More specifically, we have seen that every shellable triangulation can be transformed into a single simplex along a sequence of local bi-Lipschitz deformations, and their bi-Lipschitz constants are controlled by the shape regularity of the triangulation.

@@

\begin{remark}\label{remark:fullrecursivesum:exterior}
    \color{red}
    We provide a similar discussion as in Remark~\ref{remark:fullrecursivesum}, 
    this time geared towards the application of Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
    Recall that the potential $w$ for the gradient $\cartan u$ satisfies the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \| \cartan u      \|_{L^{p}(T_{m})} 
        \\&\qquad
        +
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \Cfivetrafo{n}{k+1}{\ell_{m}}{p}(\calT)
%         \Cfiveprime{n}{k}(\calT)^{k} 
%         \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| \cartan u \|_{L^{p}(\Makoto_{m-1})}
        \\&\qquad
        + 
        \Cfivetrafo{n}{k}{\ell_{m}}{p}(\calT) 
%         \Cfiveprime{n}{k}(\calT)^{k-1} 
%         \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(\Makoto_{m-1})}
        .
    \end{align*}
    Unwrapping the recursion for the norm over the $m$-th simplex, 
    we find an estimate of the form 
    \begin{align}\label{math:fullrecursivesum:exterior}
        \| w \|_{L^{p}(T_{m})} 
        &
        \leq 
        \sum_{\ell=0}^{M} C_{m,\ell} \| \cartan u \|_{L^{p}(T_{\ell})}
        .
    \end{align}
    Here, $C_{m,\ell}$ is the coefficient of $\| \cartan u \|_{L^{p}(T_{\ell})}$, possibly zero, 
    as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \| w \|_{L^{p}(\Omega)}^{p}
        &
        \leq 
        \sum_{m=0}^{M}
        \| w \|_{L^{p}(T_{m})}^{p}
        % \\&
        \leq 
        \sum_{m=0}^{M}
        \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
        \sum_{\ell=0}^{M} 
        \| \cartan u \|_{L^{p}(T_{\ell})}^{p} 
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
        \right)
        \| \cartan u \|_{L^{p}(\Omega)}^{p} 
        ,
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
\end{remark}

@@

Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative} and the subequent remark (using $p=2$) give an upper bound of $2/\pi \approx 0.45015815807$,
which is an overestimate by a factor of at most six.
We take this as an indication that the estimate is within feasible range, at least on simplices with good shape regularity. 
We expect the bound to deteriorate as the eccentricity of the tetrahedron increases.

However, the results indicate room for improvement. 
Future estimates should tighten the gap towards the actual computed values,
including for ill-shaped simplices.

>>>>

\begin{theorem}\label{theorem:fullrecursivesum:grad}
    Let $\calT$ be a face-connected $n$-dimensional finite triangulation
    and that the domain $\Omega$ is the interior of the underlying set of $\calT$. 
    Let $1 \leq p,q \leq \infty$ with $1 = 1/p + 1/q$,
    and suppose that $u, w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$.

    Suppose that $n$-simplices are enumerated as $T_0, T_1, \dots, T_M$,
    and that with each $n$-simplex $T_{m} \in \calT$ 
    we have a sequence of indices 
    \begin{gather*}
        0 = \imath(m,0), \; \imath(m,1), \; \dots, \; \imath(m,L_m) = m
    \end{gather*}
    such that for each $0 \leq m \leq M$ we have estimates of the following form:
    \begin{align*}
        \| w \|_{L^{p}(T_{0})} &\leq A_{m,0} \| \nabla u \|_{L^{p}(T_{0})}
        ,
    \end{align*}
    \begin{align*}
        \| w \|_{ L^{p}(T_{\imath(m,\ell)}) }
        &\leq  
        A_{m,\imath(m,\ell)}
        \| w \|_{ L^{p}(T_{\imath(m,\ell-1)}) }
        \\&\qquad 
        +
        B'_{m,\imath(m,\ell)}
        \| \nabla u \|_{ L^{p}(T_{\imath(m,\ell)}) }
        + 
        B''_{m,\imath(m,\ell)}
        \| \nabla u \|_{ L^{p}(T_{\imath(m,\ell-1)}) }
        ,
        \quad 1 \leq \ell \leq L_m
        .
    \end{align*}
    Then 
    \begin{gather*}
        \| w \|_{ L^{p}(\Omega) }
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{L_m} D_{m,\ell}^{q} \right)^{\frac p q}
        \right)^{\frac 1 p}
        \| \nabla u \|_{ L^{p}(\Omega) }
        ,
    \end{gather*}
    where 
    \begin{gather*}
        D_{m,L_m} = A_{m,m},
        \\
        D_{m,\ell} 
        = 
        \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) 
        \left( B_{m,\imath(m,\ell+1)} A_{m,\imath(m,\ell)} + A'_{m,\imath(m,\ell+1)} \right) 
        ,
        \qquad 
        1 \leq \ell \leq L_{m}-1
        .
    \end{gather*}
\end{theorem}
\begin{proof}
    Unwrapping the recursion for the norm over the $m$-th simplex, we get a total expression of the form 
    \begin{align*}%\label{math:fullrecursivesum}
        \begin{split}
            \| w \|_{L^{p}(T_{m})} 
            &
            \leq 
            \sum_{\ell=0}^{L_m}   \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+1)} \right) A_ {m,\imath(m,\ell  )} \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&\qquad 
            +
            \sum_{\ell=0}^{L_m-1} \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) A'_{m,\imath(m,\ell+1)} \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&
            = 
            A_{m,m} \| \nabla u \|_{L^{p}(T_{\imath(m,L_m)})}
            \\&\qquad 
            +
            \sum_{\ell=0}^{L_m-1} 
            \left( B_{m,\imath(m,L_m)} \cdots B_{m,\imath(m,\ell+2)} \right) 
            \left( B_{m,\imath(m,\ell+1)} A_{m,\imath(m,\ell)} + A'_{m,\imath(m,\ell+1)} \right) 
            \| \nabla u \|_{L^{p}(T_{\imath(m,\ell)})}
            \\&
            =:
            \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})}
            .
        \end{split}
    \end{align*}
    Here, $C_{m,\ell}$ is the coefficient of $\| \nabla u \|_{L^{p}(T_{\ell})}$, possibly zero, as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \begin{split}
            \| w \|_{L^{p}(\Omega)}^{p}
            &
            \leq 
            \sum_{m=0}^{M}
            \| w \|_{L^{p}(T_{m})}^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell} \| \nabla u \|_{L^{p}(T_{\ell})} \right)^{p}
            \\&
            \leq 
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \sum_{\ell'=0}^{M} \| \nabla u \|_{L^{p}(T_{\ell'})}^{p} 
            \leq 
            \left(
                \sum_{m=0}^{M}
                \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
            \right)
            \| \nabla u \|_{L^{p}(\Omega)}^{p} 
            ,
        \end{split}
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
\end{proof}

@@

    We now define, as in~\cite[Equations~{\char`\(}5.12{\char`\)} and~{\char`\(}5.14{\char`\)}]{ern2020stable} or~\cite[Equations~{\char`\(}6.7{\char`\)} and~{\char`\(}6.9{\char`\)}]{Chaum_Voh_p_rob_3D_H_curl_24}, the field $\widetilde w_{m} \in W^{p}\Alt^{k}(T_{m})$ via 
    \begin{align} \label{equation:w''_ext_der}
        {\widetilde w}_{m} := u_{|T_m} + \Xi_{1}^{\ast}\left( (w_{m-1} - u)_{|\Makoto_{m-1} } \right).
    \end{align}
    One verifies that 
    \begin{align*}
        \trace_F {\widetilde w}_{m} 
        &
        = 
        \trace_F u_{|T_m} + \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \trace_F \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \trace_F u_{|T_m} + \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \trace_F \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \trace_F \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} }
        \\&
        = 
        \trace_F w_{m-1|\Makoto_{m-1}}
        .
    \end{align*}
    Moreover, 
    \begin{align*}
        \cartan {\widetilde w}_{m} 
        &
        = 
        \cartan u_{|T_m} + \cartan \Xi_{1}^{\ast} w_{m-1|\Makoto_{m-1} } - \cartan \Xi_{1}^{\ast} u_{|\Makoto_{m-1} } 
        \\&
        = 
        \cartan u_{|T_m} + \Xi_{1}^{\ast} \cartan w_{m-1|\Makoto_{m-1} } - \Xi_{1}^{\ast} \cartan u_{|\Makoto_{m-1} } 
        = 
        \cartan u_{|T_m}
        .
    \end{align*}
    Setting $w_{m} := w_{m-1}$ over $X_{m-1}$ and $w_{m} := {\widetilde w}_{m} + w''_{m}$ over $T_{m}$, 
    we thus verify $w_{m} \in W^{p}\Alt^{k}(X_{m})$ with $\cartan w_{m} = \cartan u$ over $X_{m}$.

@@

We derive an estimate for the Poincar\'e--Friedrichs constant, in analogy to Theorem~\ref{theorem:fullrecursivesum:grad}, though the generalization to the exterior derivative faces a more complicated recursive structure. 
For ease of discussion, we choose to present the result for a general class of recursive estimates.
Of course, implementations should make use of the simplifications warranted by the specific recursive structure,
such as the recursion presented in Theorem~\ref{theorem:poincarefriedrichsestimate:exterior}.

\begin{theorem}\label{theorem:fullrecursivesum:exterior}
    % Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable $n$-dimensional triangulation $\calT$.
    Let $\calT$ be a shellable $n$-dimensional triangulation, and let the domain $\Omega \subseteq \bbR^{n}$ be the interior of the underlying set of $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling of $\calT$.
    Suppose that $u, w \in W^{p}\Alt^{k}(\Omega)$, where $1 \leq p \leq \infty$, 
    such that $\cartan w = \cartan u$ and the following recursive estimate holds:
    \begin{gather*}
        \| w \|_{L^{p}(T_{0})} 
        \leq 
        A_{0,0}
        \| \cartan u \|_{L^{p}(T_{0})}
        ,
    \end{gather*}
    and for each $T_{m} \in \calT$, $1 \leq m \leq M$ we have the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        \sum_{\ell=0}^{m} A_{m,\ell} \| \cartan u \|_{L^{p}(T_{\ell})} 
        +
        \sum_{\ell=0}^{m-1} B_{m,\ell} \| w \|_{L^{p}(T_{\ell})} 
        .
    \end{align*}
    Then we have an inequality of the form 
    \begin{align*}
        \| w \|_{L^{p}(\Omega)}
        &
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
        \right)^{\frac 1 p}
        \| \cartan u \|_{L^{p}(\Omega)}
        ,
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
    Here, 
    \begin{align*}
        C_{m,\ell} = \sum_{ m = i_L > \dots > i_1 \geq \ell } B_{i_{L},i_{L-1}} \cdots B_{i_{2},i_{1}} A_{i_{1},\ell}.
    \end{align*}
\end{theorem}
\begin{proof}    
    %     The discussion is similar as in the proof of Theorem~\ref{theorem:fullrecursivesum:grad}.
    %     this time geared towards the application of Theorem~\ref{theorem:poincarefriedrichsestimate:exterior}.
    Unwrapping the recursion, we obtain an estimate of the form 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})} 
        \leq 
        \sum_{\ell = 0}^{m}
        \underbrace{ 
            \left( 
                \sum_{ m = i_L > \dots > i_1 \geq \ell }
                B_{i_{L},i_{L-1}} \cdots B_{i_{2},i_{1}} A_{i_{1},\ell}
            \right)
        }_{ =: C_{m,\ell} }
        \| \cartan u \|_{L^{p}(T_{\ell})} 
    \end{align*}
    Here, $C_{m,\ell}$ denotes the coefficient of $\| \cartan u \|_{L^{p}(T_{\ell})}$, possibly zero, 
    as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    Once again, The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \| w \|_{L^{p}(\Omega)}^{p}
        &
        \leq 
        \sum_{m=0}^{M}
        \| w \|_{L^{p}(T_{m})}^{p}
        % \\&
        \leq 
        \sum_{m=0}^{M}
        \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
        \sum_{\ell'=0}^{M} \| \cartan u \|_{L^{p}(T_{\ell'})}^{p} 
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
        \right)
        \| \cartan u \|_{L^{p}(\Omega)}^{p} 
        ,
    \end{align*}
    where $q \in [1,\infty]$ is as described above. The proof is complete. 
    % where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
\end{proof}

\begin{remark}\label{remark:fullrecursivesum:exterior}
    Recall that the potential $w$ for the gradient $\cartan u$ satisfies the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \| \cartan u      \|_{L^{p}(T_{m})} 
        \\&\qquad
        +
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \Cfivetrafo{n}{k+1}{\ell_{m}}{p}(\calT)
%         \Cfiveprime{n}{k}(\calT)^{k} 
%         \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| \cartan u \|_{L^{p}(\Makoto_{m-1})}
        \\&\qquad
        + 
        \Cfivetrafo{n}{k}{\ell_{m}}{p}(\calT) 
%         \Cfiveprime{n}{k}(\calT)^{k-1} 
%         \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(\Makoto_{m-1})}
        .
    \end{align*}
    Unwrapping the recursion for the norm over the $m$-th simplex, 
    we find an estimate of the form 
    \begin{align}\label{math:fullrecursivesum:exterior}
        \| w \|_{L^{p}(T_{m})} 
        &
        \leq 
        \sum_{\ell=0}^{M} C_{m,\ell} \| \cartan u \|_{L^{p}(T_{\ell})}
        .
    \end{align}
    Here, $C_{m,\ell}$ is the coefficient of $\| \cartan u \|_{L^{p}(T_{\ell})}$, possibly zero, 
    as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \| w \|_{L^{p}(\Omega)}^{p}
        &
        \leq 
        \sum_{m=0}^{M}
        \| w \|_{L^{p}(T_{m})}^{p}
        % \\&
        \leq 
        \sum_{m=0}^{M}
        \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
        \sum_{\ell=0}^{M} 
        \| \cartan u \|_{L^{p}(T_{\ell})}^{p} 
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
        \right)
        \| \cartan u \|_{L^{p}(\Omega)}^{p} 
        ,
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
\end{remark}

\begin{remark}\label{remark:poincarefriedrichsestimate:exterior:contraction}
    While our main interest in this manuscript, we point out an alternative application of Proposition~\ref{proposition:starreflection},
    using the second class of transformations. 
    Those transformations are a sequence of contractions.
    Here, we take inspiration from the topological observation that domains with shellable triangulations are contractible. 
    More specifically, we have seen that every shellable triangulation can be transformed into a single simplex along a sequence of local bi-Lipschitz deformations, and their bi-Lipschitz constants are controlled by the shape regularity of the triangulation.

@@

@@

Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative} and the subequent remark (using $p=2$) give an upper bound of $2/\pi \approx 0.45015815807$,
which is an overestimate by a factor of at most six.
We take this as an indication that the estimate is within feasible range, at least on simplices with good shape regularity. 
We expect the bound to deteriorate as the eccentricity of the tetrahedron increases.
However, the results indicate room for improvement. 
Future estimates should tighten the gap towards the actual computed values,
including for ill-shaped simplices.
